Simple and robust inference of connectomes using partial correlation coefficients

### 2014-06-09 article.tex
Fix stacking notation

%\paragraph{Motivation\\} Intuitively, connectivity information between two
neurons may be retrieved from successive and slightly delayed burst events. In
case of a directed relationship, the burst order should be the same while in
case of an undirected relationship, it could change. Therefore, the
fluorescent signal can be reduced to a binary signal representing the activity
of a neuron. This is the idea behind the filter $g_1$:  \begin{align}
g_1(X^i_t) = (X^i_t)^{0.9} \mathbb{1}(X^i_t \ge h) \end{align} where ... .

Considering the short delay for a communication between neurons, slow events
can be neglected to retrieve direct relationships. Note that the average time
delay is about $1ms$ or $2ms$ while the fluorescent signal time resolution is
$20ms$. Let us also highlight that even if two neurons are spatially far from
each other, the information is expected in the same time step or at most in
the next time step. It is therefore not necessary to keep low-frequency event
and we thus apply a high-pass filter $g_2$: \begin{align} g_2(X^i_t) =
\sum_{l=-3}^{2} c^k_l X^i_{t+l}  \end{align} where ... .

@@

@@

\begin{align}
g_4(X_t^i) = (X_t^i + 1)^{\left( 1 + \dfrac{1}{a_t} \right)} &\text{ with } a_t = \sum_{i=1}^p \left ( X_t^i + \dfrac{1}{2} X_{t-1}^i \right ) \\
g_5(X_t^i) = (X_t^i)^{h(r_t)} &\text{ with } r_t = \frac{a_t}{\max_{l \in T}{a_l}} \\
\end{align}

where $h(x)$ is a 4-piecewise continuous functions whose parameters were tuned
for each low-pass filter.

Hence, we define our preprocessing function $G$ as 
\begin{align}
G(X_t^i) = g_5 \bullet g_4 \bullet g_1 \bullet g_3 \bullet g_2(X_t^i).
\end{align}

@@

Then we apply on all remaining timesteps $T$ a non-linear filter based on the current
activity of the systems in order to favor the association measure when
a medium size set of neurons are firing together, while artificially
reducing the association measure in case of extreme network activity,
i.e. high when a network firing is happening and low when no neuron is
active:
\begin{align}
a_t &= \sum_{i=1}^p X^{\prime\prime\prime}_{t,i} +
                   \frac{1}{2} X^{\prime\prime\prime}_{t-1,i} \forall t \in T\\
r_t &= \frac{a_t}{\max_{l \in T}{a_l}} \\
X^{\prime\prime\prime\prime}_{t,i} &=
\left\{
  \begin{array}{l}
    1  : r_t < 0\\
    {(X^{\prime\prime\prime}_{t,i} + 1)^{(a_t^{-1} + 1)}}^{g(r_t)} : r_t \geq 0
  \end{array}
\right.
\end{align}
where $g(x) = b_1 I(f_4 < x < f_5) + b_2  I(f_5 \leq x < f_6) +
b_3 (I(x\leq f_4) + I(f_ 6 \leq x))$ is a piecewise continuous function of
parameters $b_1, b_2, b_3, f_4, f_5, f_6$. The coefficient were tuned
for each set of coefficient $\left\{c_l^k\right\}$.
\hrule

@@

% Speak about why averaging
The quality of the signal filtering and network inference is dependent upon
good hyper-parameter selection, which are themselves dependent on the properties
of the targeted network and measurement conditions, e.g. signal to noise ratio.
Ensemble methods are known to improve robustness with respect to parameter
selection [todo cite] and improve performance [todo cite].
A weighted average $y_{stack}$ of partial correlation matrices was made
by varying uniformly the hard-threshold value (parameter $f_3$) and the design
of the low-pass filter (coefficients $c_l$). The stacked connectivity score
matrix $y_{stacking}$ is given by
\begin{align}
y_{stack} =  \dfrac{1}{\sum_{k\in\{1,..,4\}} w_k |T_k|} \sum_{k=1}^{k=4}  w_k \sum_{f_3 \in T_k} f(X, k, f_3)
\end{align}
where $f(X, k, f_3)$ gives the predicted adjacency matrix, i.e. applies the
filtering and inference methodologies (as described
in sections \ref{sec:filter} and \ref{sec:inference}), with a  given value
$f_3$ for the hard-threshold among a set of thresholds $T_k$ of size $|T_k|$
and the $k^{th}$ low-pass filter pattern associated to a weight $w_k$ whose
values are $w_1 = 0.01 , w_2 =1, w_3 = 0.7, w_4 = 0.9$.

>>>>

%\paragraph{Motivation\\} Intuitively, connectivity information between two
neurons may be retrieved from successive and slightly delayed burst events. In
case of a directed relationship, the burst order should be the same while in
case of an undirected relationship, it could change. Therefore, the
fluorescent signal can be reduced to a binary signal representing the activity
of a neuron. This is the idea behind the filter $g_1$:  \begin{align}
g_1^q(X^i_t) = (X^i_t)^{0.9} \mathbb{1}(X^i_t \ge q) \end{align} where ... .

Considering the short delay for a communication between neurons, slow events
can be neglected to retrieve direct relationships. Note that the average time
delay is about $1ms$ or $2ms$ while the fluorescent signal time resolution is
$20ms$. Let us also highlight that even if two neurons are spatially far from
each other, the information is expected in the same time step or at most in
the next time step. It is therefore not necessary to keep low-frequency event
and we thus apply a high-pass filter $g_2$: \begin{align} g_2^k(X^i_t) =
\sum_{l=-3}^{2} c^k_l X^i_{t+l}  \end{align} where ... .

@@

% We then define our preprocessing function $G$ as
% \begin{align}
% G(X_t^i) = g^q_1(g_3(g_2(X_t^i))).
% \end{align}

@@

\begin{align}
h_1(X_t^i) = (X_t^i + 1)^{\left( 1 + \dfrac{1}{a_t} \right)} &\text{ with } a_t = \sum_{i=1}^p \left ( X_t^i + \dfrac{1}{2} X_{t-1}^i \right ) \\
h_2(X_t^i) = (X_t^i)^{r(a_t)} &\text{ with } r_t = \frac{a_t}{\max_{l \in T}{a_l}} \\
\end{align}

where $r(x)$ is a 4-piecewise continuous functions whose parameters were tuned
for each low-pass filter.

% Hence, we define our preprocessing function $H$ as 
% \begin{align}
% H(X_t^i) = h_2 \bullet h_1 \bullet G(X).
% \end{align}

Hence, we define our preprocessing function $G$ as 
\begin{align}
G^{q,k}(X_t^i) = h_2 \bullet h_1 \bullet g^q_1 \bullet g_3 \bullet g_2^k(X_t^i).
\end{align}

@@

Then we apply on all remaining timesteps $T$ a non-linear filter based on the current
activity of the systems in order to favor the association measure when
a medium size set of neurons are firing together, while artificially
reducing the association measure in case of extreme network activity,
i.e. high when a network firing is happening and low when no neuron is
active:
\begin{align}
a_t &= \sum_{i=1}^p X^{\prime\prime\prime}_{t,i} +
                   \frac{1}{2} X^{\prime\prime\prime}_{t-1,i} \forall t \in T\\
r_t &= \frac{a_t}{\max_{l \in T}{a_l}} \\
X^{\prime\prime\prime\prime}_{t,i} &=
\left\{
  \begin{array}{l}
    1  : r_t < 0\\
    {(X^{\prime\prime\prime}_{t,i} + 1)^{(a_t^{-1} + 1)}}^{g(r_t)} : r_t \geq 0
  \end{array}
\right.
\end{align}
where $g(x) = b_1 I(f_4 < x < f_5) + b_2  I(f_5 \leq x < f_6) +
b_3 (I(x\leq f_4) + I(f_ 6 \leq x))$ is a piecewise continuous function of
parameters $b_1, b_2, b_3, f_4, f_5, f_6$. The coefficient were tuned
for each set of coefficient $\left\{c_l^k\right\}$.

@@

% Speak about why averaging
The quality of the signal filtering and network inference is dependent upon
good hyper-parameter selection, which are themselves dependent on the properties
of the targeted network and measurement conditions, e.g. signal to noise ratio.
Ensemble methods are known to improve robustness with respect to parameter
selection [todo cite] and improve performance [todo cite].
A weighted average $y_{stack}$ of partial correlation matrices was made
by varying uniformly the hard-threshold value (parameter $f_3$) and the design
of the low-pass filter (coefficients $c_l$). The stacked connectivity score
matrix $y_{stacking}$ is given by
\begin{align}
y_{stack} =  \dfrac{1}{\sum_{k\in\{1,..,4\}} w_k |T_k|} \sum_{k=1}^{k=4}  w_k \sum_{q \in T_k} G^{q,k}(X_t^i)
\end{align}
where $f(X, k, f_3)$ gives the predicted adjacency matrix, i.e. applies the
filtering and inference methodologies (as described
in sections \ref{sec:filter} and \ref{sec:inference}), with a  given value
$f_3$ for the hard-threshold among a set of thresholds $T_k$ of size $|T_k|$
and the $k^{th}$ low-pass filter pattern associated to a weight $w_k$ whose
values are $w_1 = 0.01 , w_2 =1, w_3 = 0.7, w_4 = 0.9$.
